An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-31 VLDB2026/main.tex
Update main.tex

	\begin{figure*}
		\centering
		\setlength{\belowcaptionskip}{-0.cm}
		\includegraphics[width=0.92\textwidth]{figures/hybrid ANNS.pdf}
		\caption{Hybrid Query Example}

@@

	\textcolor{violet}{IVF-based indexes also offer unique advantages in hybrid ANN search scenarios, including natural support for pre-filtering, low memory footprint, and small index size. Efficient filtering is achieved by excluding vectors that do not satisfy attribute constraints before distance computation (e.g., Faiss, Milvus). Furthermore, hierarchical partitioning by attributes can form finer-grained sub-clusters to enhance filtering efficiency (e.g., CAPS, Puck).}
%	IVF-based ANN search algorithms improve computational efficiency by partitioning the dataset into clusters and restricting the search to clusters nearest to the query point. Specifically, as shown in Figure~\ref{fig:ivf}, IVF first selects a subset of data points and applies a clustering algorithm (e.g., K-Means) to obtain a set of centroids $(C_0, C_1, ..., C_6)$, each representing a distinct cluster. Every data point (gray dots) is then assigned to the cluster corresponding to its nearest centroid based on the distance metric, thereby forming the index structure.
%	
%	During the query process, given a query point $q$, the algorithm computes distances between $q$ and all centroids, selects a small number of the closest centroids ($C_0, C_1, C_2$), and then performs a search only within the associated clusters to identify the approximate nearest neighbors.
%	
%	\textcolor{violet}{IVF-based indexes also offer unique advantages in Hybrid ANN search scenarios, such as naturally supporting pre-filtering and requiring significantly lower memory consumption and smaller index sizes compared to graph-based indexes. With minor modifications, IVF indexes achieve attribute and range filtering by directly excluding vectors that do not meet query constraints through pre-filtering, thus skipping irrelevant vectors during intra-cluster distance computations (e.g., Faiss, Milvus). Moreover, clusters can be further partitioned hierarchically based on attributes (hierarchical partitioning), resulting in finer-grained sub-partitions that significantly enhance the efficiency of attribute filtering (e.g., CAPS, PUCK).}

@@

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
%				\textit{AND} and \textit{OR} indicate whether the algorithm supports the corresponding logical operations.
%				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
%				\textit{Complex Boolean} indicates if the algorithm supports complex expressions, such as using the NOT operator or nested logic with parentheses.
%				\textit{Dynamic Insert} indicates whether the algorithm supports the operation of dynamically inserting data points.
%				\textit{Multi Thread} indicates whether the algorithm supports multi-threaded search.
%				\textit{Disk Support} indicates whether the algorithm supports disk storage.
%				\textit{Base} indicates the foundational ANN algorithm for the hybrid query method.
		\textcolor{orange}{
		\textit{Filtered}: FilteredVamana, \textit{Stitched}: StitchedVamana, \textit{Win}: WinFilter.
		\textit{AND} / \textit{OR}: Whether the algorithm supports the corresponding logical operations.
		\textit{Flexible Attributes}: Allows the number of query attributes to differ from that in index construction.
		\textit{Complex Boolean}: Supports NOT or nested Boolean expressions.
		\textit{Dynamic Insert}: Supports dynamic data insertion.
		\textit{Multi Thread}: Supports multi-threaded search.
		\textit{Disk Support}: Supports disk-based storage.
		\textit{Base}: Underlying ANN index structure used.}
		\textit{A}: Post-filtering (Filter after ANN search), 
		\textit{B}: Pre-filtering (Filter before ANN search), 
		\textit{C}: Simultaneous filtering (Filter simultaneously with ANN search), 
		\textit{Y}: Support, 
		\textit{N}: Unsupport. 
		\end{minipage}}

@@

		%		\caption{Time and space overhead of attribute filtering index construction}
		\caption{\textcolor{violet}{Time and Space Overhead (AF)}}
		\label{fig:build_index_comparison}
	\end{figure}

@@

		\includegraphics[width=0.95\textwidth]{figures/exp/exp_1_1_SingleLabel_1thread.pdf}
		\caption{Performance of Single-Attribute Index Construction and Single-Attribute Query }
		\label{fig:exp_1_1_SingleLabel_1thread}
	\end{figure*}

@@

		% 上面的图例，居中并可通过 hspace 调整左右位置
		\hspace*{15pt} % 可调整的参数，负值左移，正值右移，例如 -20pt 或 20pt
		\includegraphics[width=0.85\columnwidth]{figures/exp/exp_2_legend.pdf} % 图例图片路径

@@

		\includegraphics[width=0.95\textwidth]{figures/exp/exp_4_1_MultiLabel_1thread.pdf}
		\caption{Performance of Multi-Attribute Index Construction and Multi-Attribute Query}
		\label{fig:exp_4_1_MultiLabel_1thread}
	\end{figure*}

	\includegraphics[width=0.95\textwidth]{figures/exp/exp_3_1.pdf}
	\caption{
		Effect of Attribute Distribution on Query Performance}
%		(Strong is robust to attribute distribution shifts, Weak is sensitive to attribute distribution shifts.)}

@@

	\begin{minipage}[t]{0.38\textwidth}
		\centering
		% 内部图片宽度使用\linewidth，它会自适应为容器宽度(0.32\textwidth)
		\includegraphics[width=0.495\linewidth]{figures/exp/attribute_multimodel.pdf}
		\hfill 
		\includegraphics[width=0.47\linewidth]{figures/exp/attribute_multimodel_1.pdf}

	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.203\textwidth}
		\centering
		\includegraphics[width=0.96\linewidth]{figures/exp/attribute_100M.pdf}

	\end{minipage}
	\hfill % 在盒子之间添加弹性间距
	\begin{minipage}[t]{0.39\textwidth}
		\centering
		\includegraphics[width=0.495\linewidth]{figures/exp/attribute_85.pdf}
		\hfill
		\includegraphics[width=0.47\linewidth]{figures/exp/attribute_71.pdf}

@@

		%		\caption{Time and space overhead of range filtering index construction}
%		\caption{Time and space overhead of index construction}
		\caption{\textcolor{violet}{Time and Space Overhead (RF)}}
		\label{fig:rangeFilter_build_index_comparison}
	\end{figure}
	\subsection{Range Filtering}
	\textcolor{blue}{Due to the limitations of existing range filtering algorithms, our experiment uses only a single attribute, with the dataset sorted in ascending order by value. The WIT-Real and YT-Audio-Real datasets use real attribute values, while the rest use randomly generated ones. }
	\subsubsection{\textcolor{violet}{Time and Space overhead.}}

	%Similar to the evaluation of \textit{Attribute Filtering} algorithms, 
	Similarly, we also evaluate \textcolor{orange}{4} key metrics mentioned above in \textit{Range Filtering} algorithms. \textcolor{orange}{Notably, all algorithms use single-threaded index construction on medium-scale datasets, and 64-thread construction on the large-scale Deep100M dataset, except DSG, which only supports single-threading. For WinFilter, we use its integrated Super-Postfiltering on medium datasets for best performance, and switch to its base algorithm Vamana-WST on Deep100M due to construction overhead.
	}

	\textit{\textbf{Index construction time.}}
%	As shown in the Figure~\ref{fig:rangeFilter_build_time}, for regular-scale datasets, Faiss builds the index fastest. As Faiss obtains subsamples through sampling and calculates the cluster centroid, the computational complexity is significantly reduced.
%
%	WinFilter and ACORN-\(\gamma\) take the longest.
%	WinFilter relies on a tree structure, requiring separate nearest-neighbor graphs per node. This causes redundant computation and high costs, especially for large datasets.
%	ACORN-\(\gamma\) expands the neighbor list of each node to provide more candidate paths, which increases both computation and storage costs. It also prunes distant neighbors to save storage space. However, this introduces additional overhead and slows down the construction process.
%	
%	\textcolor{orange}{On large-scale datasets, ACORN-1 achieves the fastest index construction, primarily due to its similarity to the original HNSW construction process and its use of fewer candidate neighbors in the upper layers. SeRF, another HNSW-based method, also exhibits very short construction time by losslessly compressing multiple HNSW graphs into a unified structure and accelerating the process through dynamic pruning and skip-connection optimizations.
%	As two methods both based on IVF\_Flat, Milvus and Faiss achieve short indexing times. Milvus is even faster, mainly due to its cache-aware optimization, which significantly reduces CPU cache misses.}
%	\textcolor{orange}{
%	DSG only supports single-threaded construction, resulting in the longest build time.}
	\textcolor{orange}{
	As shown in Figure~\ref{fig:rangeFilter_build_time}, Faiss and Milvus, based on the IVF architecture, require less time. This indexing structure enables fast construction through subset sampling and centroid computation, thereby reducing computational cost. Milvus performs particularly well on large-scale datasets, mainly due to its cache-aware optimizations, which significantly reduce CPU cache misses.}
%	As shown in Figure~\ref{fig:rangeFilter_build_time}, Faiss, based on the IVF\_Flat architecture, achieves fast index construction through subset sampling and centroid computation, which lowers the computational cost. 
%	As another IVF\_Flat-based method, Milvus performs even better on large-scale datasets, mainly due to its cache-aware optimization that greatly reduces CPU cache misses.}

@@

	In this experiment, we follow the query range definition adopted by prior work~\cite{HQI}. \textcolor{red}{Specifically, for a given query, the range ratio is defined as $2^{-i}$	when the query range covers $\frac{n}{2^i}$ data points, where $n$ is the total dataset size.}
	% Based on this definition, we evaluate algorithm performance under 4 query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. \textcolor{blue}{Due to the limitations of existing range filtering algorithms, the range filtering experiment only uses one attribute, and the dataset is sorted in ascending order according to the attribute value size. WIT-Real and YT-Audio-Real use real attribute values, and the rest use randomly generated attribute values.}
	%\textcolor{red}{Some algorithms don't support using actual attributes as search keys, instead requiring an ID (the primary key, ranging from $0$ to $n-1$). Therefore, we uniformly use this ID as the search key. This ID is pre-sorted according to its corresponding real attribute, allowing us to map an attribute to its ID without loss of generality for efficient range queries.} 
	Based on this definition, we evaluate algorithm performance under the largest ($2^{-2}$) and smallest ($2^{-8}$) query range settings. We report the results for these two representative cases in Figure~\ref{fig:exp_8_2}.
%	The experimental results show consistent performance across all datasets and query ranges.

@@

	\textit{\textbf{\textcolor{red}{Large-scale Dataset.}}} \textcolor{red}{As shown in the Figure~\ref{fig: range big dataset}, On the Deep100M dataset, the top 3 ranked algorithms (WinFilter, UNIFY, and iRange) remain unchanged. However, their high performance is achieved at the cost of extremely high memory consumption (Figure~\ref{fig:rangeFilter_search_memory_mb}). In contrast, VBASE and Faiss demonstrate moderate performance while consuming significantly less memory.}

>>>>

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.2cm}
		\setlength{\belowcaptionskip}{-0.1cm}

		\includegraphics[width=0.92\textwidth]{figures/hybrid ANNS.pdf}
		\caption{Hybrid Query Example}

@@

	\textcolor{violet}{IVF-based indexes also offer unique advantages in hybrid ANN search scenarios, including natural support for pre-filtering, low memory footprint, and small index size. Efficient filtering is achieved by excluding vectors that do not satisfy attribute constraints before distance computation (e.g., Faiss, Milvus). Furthermore, hierarchical partitioning by attributes can form finer-grained sub-clusters to enhance filtering efficiency (e.g., CAPS, PUCK).}
%	IVF-based ANN search algorithms improve computational efficiency by partitioning the dataset into clusters and restricting the search to clusters nearest to the query point. Specifically, as shown in Figure~\ref{fig:ivf}, IVF first selects a subset of data points and applies a clustering algorithm (e.g., K-Means) to obtain a set of centroids $(C_0, C_1, ..., C_6)$, each representing a distinct cluster. Every data point (gray dots) is then assigned to the cluster corresponding to its nearest centroid based on the distance metric, thereby forming the index structure.
%	
%	During the query process, given a query point $q$, the algorithm computes distances between $q$ and all centroids, selects a small number of the closest centroids ($C_0, C_1, C_2$), and then performs a search only within the associated clusters to identify the approximate nearest neighbors.
%	
%	\textcolor{violet}{IVF-based indexes also offer unique advantages in Hybrid ANN search scenarios, such as naturally supporting pre-filtering and requiring significantly lower memory consumption and smaller index sizes compared to graph-based indexes. With minor modifications, IVF indexes achieve attribute and range filtering by directly excluding vectors that do not meet query constraints through pre-filtering, thus skipping irrelevant vectors during intra-cluster distance computations (e.g., Faiss, Milvus). Moreover, clusters can be further partitioned hierarchically based on attributes (hierarchical partitioning), resulting in finer-grained sub-partitions that significantly enhance the efficiency of attribute filtering (e.g., CAPS, PUCK).}

@@

		\centering
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
%				\textit{AND} and \textit{OR} indicate whether the algorithm supports the corresponding logical operations.
%				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
%				\textit{Complex Boolean} indicates if the algorithm supports complex expressions, such as using the NOT operator or nested logic with parentheses.
%				\textit{Dynamic Insert} indicates whether the algorithm supports the operation of dynamically inserting data points.
%				\textit{Multi Thread} indicates whether the algorithm supports multi-threaded search.
%				\textit{Disk Support} indicates whether the algorithm supports disk storage.
%				\textit{Base} indicates the foundational ANN algorithm for the hybrid query method.
		\textcolor{orange}{
		\textit{Filtered}: FilteredVamana, \textit{Stitched}: StitchedVamana, \textit{Win}: WinFilter.
		\textit{AND} / \textit{OR}: Whether the algorithm supports the corresponding logical operations.
		\textit{Flexible Attributes}: Allows the number of query attributes to differ from that in index construction.
		\textit{Complex Boolean}: Supports NOT or nested Boolean expressions.
		\textit{Dynamic Insert}: Supports dynamic data insertion.
		\textit{Multi Thread}: Supports multi-threaded search.
		\textit{Disk Support}: Supports disk-based storage.
		\textit{Base}: Underlying ANN index structure used.}
		\textit{A}: Post-filtering (Filter after ANN search), 
		\textit{B}: Pre-filtering (Filter before ANN search), 
		\textit{C}: Simultaneous filtering (Filter simultaneously with ANN search), 
		\textit{Y}: Support, 
		\textit{N}: Unsupport. 
		\end{minipage}}

@@

		%		\caption{Time and space overhead of attribute filtering index construction}
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.3cm}
		\caption{\textcolor{violet}{Time and Space Overhead (AF)}}
		\label{fig:build_index_comparison}
	\end{figure}

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.2cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_1_1_SingleLabel_1thread.pdf}
		\caption{Performance of Single-Attribute Index Construction and Single-Attribute Query }
		\label{fig:exp_1_1_SingleLabel_1thread}
	\end{figure*}

@@

	\begin{figure}[t]
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.3cm}
		% 上面的图例，居中并可通过 hspace 调整左右位置
		\hspace*{15pt} % 可调整的参数，负值左移，正值右移，例如 -20pt 或 20pt
		\includegraphics[width=0.85\columnwidth]{figures/exp/exp_2_legend.pdf} % 图例图片路径

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.2cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_4_1_MultiLabel_1thread.pdf}
		\caption{Performance of Multi-Attribute Index Construction and Multi-Attribute Query}
		\label{fig:exp_4_1_MultiLabel_1thread}
	\end{figure*}

	\begin{figure*}
	\centering
	\setlength{\abovecaptionskip}{0.1cm}
	\setlength{\belowcaptionskip}{-0.2cm}
	\includegraphics[width=0.95\textwidth]{figures/exp/exp_3_1.pdf}
	\caption{
		Effect of Attribute Distribution on Query Performance}
%		(Strong is robust to attribute distribution shifts, Weak is sensitive to attribute distribution shifts.)}

@@

	\begin{minipage}[t]{0.38\textwidth}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.2cm}
		% 内部图片宽度使用\linewidth，它会自适应为容器宽度(0.32\textwidth)
		\includegraphics[width=0.495\linewidth]{figures/exp/attribute_multimodel.pdf}
		\hfill 
		\includegraphics[width=0.47\linewidth]{figures/exp/attribute_multimodel_1.pdf}

	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.203\textwidth}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.2cm}
		\includegraphics[width=0.96\linewidth]{figures/exp/attribute_100M.pdf}

	\end{minipage}
	\hfill % 在盒子之间添加弹性间距
	\begin{minipage}[t]{0.39\textwidth}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.2cm}
		\includegraphics[width=0.495\linewidth]{figures/exp/attribute_85.pdf}
		\hfill
		\includegraphics[width=0.47\linewidth]{figures/exp/attribute_71.pdf}

@@

		%		\caption{Time and space overhead of range filtering index construction}
%		\caption{Time and space overhead of index construction}
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.3cm}
		\caption{\textcolor{violet}{Time and Space Overhead (RF)}}
		\label{fig:rangeFilter_build_index_comparison}
	\end{figure}
	\subsection{Range Filtering}
	\textcolor{blue}{Due to the limitations of existing range filtering algorithms, our experiment uses only a single attribute, with the dataset sorted in ascending order by value. The WIT-Real and YT-Audio-Real datasets use real attribute values, while the rest use randomly generated ones. }
	\subsubsection{\textcolor{violet}{Time and Space overhead.}}

	%Similar to the evaluation of \textit{Attribute Filtering} algorithms, 
	Similarly, we also evaluate \textcolor{orange}{4} key metrics mentioned above in \textit{Range Filtering} algorithms. \textcolor{orange}{Notably, all algorithms use single-threaded index construction on medium-scale datasets, and 64-thread construction on the large-scale Deep100M dataset, except DSG, which only supports single-threading. For WinFilter, we use Super Postfiltering on medium datasets for best performance, and switch to the base algorithm Vamana-WST on Deep100M due to construction overhead.
	}

	\textit{\textbf{Index construction time.}}
%	As shown in the Figure~\ref{fig:rangeFilter_build_time}, for regular-scale datasets, Faiss builds the index fastest. As Faiss obtains subsamples through sampling and calculates the cluster centroid, the computational complexity is significantly reduced.
%
%	WinFilter and ACORN-\(\gamma\) take the longest.
%	WinFilter relies on a tree structure, requiring separate nearest-neighbor graphs per node. This causes redundant computation and high costs, especially for large datasets.
%	ACORN-\(\gamma\) expands the neighbor list of each node to provide more candidate paths, which increases both computation and storage costs. It also prunes distant neighbors to save storage space. However, this introduces additional overhead and slows down the construction process.
%	
%	\textcolor{orange}{On large-scale datasets, ACORN-1 achieves the fastest index construction, primarily due to its similarity to the original HNSW construction process and its use of fewer candidate neighbors in the upper layers. SeRF, another HNSW-based method, also exhibits very short construction time by losslessly compressing multiple HNSW graphs into a unified structure and accelerating the process through dynamic pruning and skip-connection optimizations.
%	As two methods both based on IVF\_Flat, Milvus and Faiss achieve short indexing times. Milvus is even faster, mainly due to its cache-aware optimization, which significantly reduces CPU cache misses.}
%	\textcolor{orange}{
%	DSG only supports single-threaded construction, resulting in the longest build time.}
	\textcolor{orange}{
	As shown in Figure~\ref{fig:rangeFilter_build_time}, Faiss and Milvus, based on the IVF\_Flat architecture, require less time. This indexing structure enables fast construction through subset sampling and centroid computation, thereby reducing computational cost. Milvus performs particularly well on large-scale datasets, mainly due to its cache-aware optimizations, which significantly reduce CPU cache misses.}
%	As shown in Figure~\ref{fig:rangeFilter_build_time}, Faiss, based on the IVF\_Flat architecture, achieves fast index construction through subset sampling and centroid computation, which lowers the computational cost. 
%	As another IVF\_Flat-based method, Milvus performs even better on large-scale datasets, mainly due to its cache-aware optimization that greatly reduces CPU cache misses.}

@@

	In this experiment, we follow the query range definition adopted by prior work~\cite{HQI}. \textcolor{red}{Specifically, for a given query, the range ratio is defined as $2^{-i}$	when the query range covers $\frac{n}{2^i}$ data points, where $n$ is the total dataset size.}
	% Based on this definition, we evaluate algorithm performance under 4 query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. \textcolor{blue}{Due to the limitations of existing range filtering algorithms, the range filtering experiment only uses one attribute, and the dataset is sorted in ascending order according to the attribute value size. WIT-Real and YT-Audio-Real use real attribute values, and the rest use randomly generated attribute values.}
	%\textcolor{red}{Some algorithms don't support using actual attributes as search keys, instead requiring an ID (the primary key, ranging from $0$ to $n-1$). Therefore, we uniformly use this ID as the search key. This ID is pre-sorted according to its corresponding real attribute, allowing us to map an attribute to its ID without loss of generality for efficient range queries.} 
	Based on this definition, we evaluate algorithm performance under four query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. For space considerations, Figure~\ref{fig:exp_8_2} reports results for only the largest ($2^{-2}$) and smallest ($2^{-8}$) range ratios. 
%	The experimental results show consistent performance across all datasets and query ranges.

@@

	\textit{\textbf{\textcolor{red}{Large-scale Dataset.}}} \textcolor{red}{As shown in the Figure~\ref{fig: range big dataset}, On the deep100M dataset, the top 3 ranked algorithms (WinFilter, UNIFY, and iRange) remain unchanged. However, their high performance is achieved at the cost of extremely high memory consumption (Figure~\ref{fig:rangeFilter_search_memory_mb}). In contrast, VBASE and Faiss demonstrate moderate performance while consuming significantly less memory.}
